---
output: html_document
editor_options: 
  chunk_output_type: console
---
---
title: "Finfish Variance Decomposition"
author: "Jack Elstner"
date: "2/27/2023"
output: pdf_document

In this document, I use Bayesian multilevel models to perform a variance decomposition on the Poseidon finfish estimates. This technique allows me to partition total variance in length estimtates into different components (user variance, species variance). This will allow me to identify factors that explain the most significant portion of the variation in Poseidion performance. I am most interested in calculating relative variance (percentages)

Questions:

1. Is there a common observation error, or does observation error differ by humans?
2. Is there a common observation error, or does observation error differ by species?


This chunk of code reads in the data and performs necessary data cleaning.
```{r}
library(rethinking)

# read in cleaned finfish data
finfish.all <- read.csv("~/Desktop/SIO/Research/Poseidon Project/Final Analysis/Final Data/Fishery-Specific Data/CaliforniaFinfish_March2_Merged&Cleaned.csv")

finfish <- subset(finfish.all, Include == "yes")

# read in angler hand meaasurements
AnglerHandMeasurements <- read.csv("~/Desktop/SIO/Research/Poseidon Project/Final Analysis/Final Data/Fishery-Specific Data/AnglerHandMeasurements.csv")

finfish <- merge(finfish, AnglerHandMeasurements, by = c("AnglerID", "TagID"))

# add fish species index variable
finfish$SpeciesID <- as.integer(as.factor(finfish$Species))

# add FishID.b index variable
finfish$FishID.b <- as.integer(as.factor(finfish$TagID))

# create AnglerNumber variable that reassigns an angler number giving angler a number 1 through 9
finfish$AnglerNumber <- as.integer(as.factor(finfish$AnglerID))
```



Model Version 1: Modeling Unbiased Variance Across Measurement Types
```{r}
# make z-score function
zScore <- function(x) {
  result <- (x - mean(x))/sd(x)
  return(result)
}

# create data list with z-scored data
datZ <- list(
  B = zScore(finfish$Angler_Hand_Measurement_TL_mm),
  O = zScore(finfish$Algorithm_Estimate_mm), 
  A = zScore(finfish$Adjusted_Estimate_mm),
  FishID = finfish$FishID.b 
)

# fit model

# TrueSize[FishID] is the true length of any given fish
# B is Board Measurements

M1 <- ulam(
  alist(
    TrueSize[FishID] ~ dnorm(0, 2),
    B ~ dnorm(TrueSize[FishID], Bsig), # board measurements are distributed without bias around some true measurement of a fish 
    O ~ dnorm(TrueSize[FishID], Osig), # original Poseidon estimates are distributed without bias around some true measurement of a fish
    A ~ dnorm(TrueSize[FishID], Asig), # adjusted Poseidon estimates are distrubuted without bias around some true measurement of a fish
    Bsig ~ dexp(1),
    Osig ~ dexp(1),
    Asig ~ dexp(1)
  ), data = datZ, chains = 4 
)

# create a traceplot of model 1 to assess convergence
traceplot(M1)

# extract samples
PostM1 <- extract.samples(M1, n = 10000)

mean(PostM1$Bsig)*

###### Figure Generation
# load packages
library(ggplot2)
library(reshape)

M1.var <- data.frame(
  PostM1$Osig,
  PostM1$Asig,
  PostM1$Bsig
)

Names <- c("Original Estimate", "Adjusted Estimates", "Hand Measurements")
names(M1.var) <- Names
M1.var <- melt(M1.var)
names2 <- c("Measurement", "Variance")
names(M1.var) <- names2

cols <- c("gray", "red", "blue")

# create density plots of variance estimates across fisheries
M1.var_plot = ggplot(data = M1.var, aes(x=Variance, fill=Measurement)) +
  geom_density(alpha=.25) +
  ggtitle("Length Measurement Variance Across Measurement Types") +
  labs(y= "Density", x = "Posterior Variance Estimates (z-score)") +
  scale_fill_manual(values = cols)

M1.var_plot + theme(plot.title = element_text(hjust = 0.5))


# calculate MAP and 95% HPDI for each posterior variance estimate
# MAP
round(chainmode(PostM1$Bsig), digits = 3)*sd(finfish$Angler_Hand_Measurement_TL_mm) # Bsig = 0.156
round(chainmode(PostM1$Osig), digits = 3)*sd(finfish$Angler_Hand_Measurement_TL_mm) # Osig = 0.613
round(chainmode(PostM1$Asig), digits = 3)*sd(finfish$Angler_Hand_Measurement_TL_mm) # Asig = 0.242

# 95% HPDI
round(HPDI(PostM1$Bsig, prob = 0.95), digits = 3) # Bsig = [0.142, 0.169] 
round(HPDI(PostM1$Osig, prob = 0.95), digits = 3) # Osig = [0.563, 0.665]
round(HPDI(PostM1$Asig, prob = 0.95), digits = 3) # Asig = [0.221, 0.262]


# posterior predictive check???

# extract samples
PostM1 <- extract.samples(M1, n = 10000)

# compute mean posterior estimate of true size for each fish
PostMeanTS <- apply(PostM1$TrueSize, 2, mean)

# Compute mean observed hand measurement for each fish
obs <- data.frame(finfish$FishID.b, zScore(finfish$Scientist_Hand_Measured_TL_mm))

ObsMeanTS <- apply(obs$zScore.finfish.Scientist_Hand_Measured_TL_mm.[obs$zScore.finfish.Scientist_Hand_Measured_TL_mm.], 2, mean)

plot(meanTS)
points(zScore(finfish$Scientist_Hand_Measured_TL_mm), col="red")
```


In this code chunk, we extract the modeled true size estimates from Model 1 for each fish and use these values to re-calculate total error and percent error metrics for finfish
```{r}
# here, we create a vector with the model-derived length estimates for each fish
# then, we add model.derived estimates to main finfish data frame
ModelLengths.z <- numeric()
for(i in 1:40) {
  ModelLengths.z <- c(ModelLengths.z, (round(chainmode(PostM1$TrueSize[ , i]), digits = 2)))
}

# transform from z-score to length measurements in mm
ModelLengths.mm <- (ModelLengths.z*sd(finfish$Angler_Hand_Measurement_TL_mm)) + mean(finfish$Angler_Hand_Measurement_TL_mm)
ModelLengths.mm
# make data frame with FishID.b and model-derived True Size Estimates
FishID.b <- 1:40
ModeledLengths.df <- data.frame(FishID.b, ModelLengths.mm)

# merge modeled lengths estimates into main finfish data frame
finfish <- merge(finfish, ModeledLengths.df, by = "FishID.b")

# re-calculate Total Error and Percent Error Metrics
finfish$New_Algorithm_Total_Error_mm <- finfish$Algorithm_Estimate_mm - finfish$ModelLengths.mm
finfish$New_Adjusted_Total_Error_mm <- finfish$Adjusted_Estimate_mm - finfish$ModelLengths.mm
finfish$New_Algorithm_Percent_Error <- (finfish$New_Algorithm_Total_Error_mm/finfish$ModelLengths.mm)*100
finfish$New_Adjusted_Percent_Error <- (finfish$New_Adjusted_Total_Error_mm/finfish$ModelLengths.mm)*100

write.csv(finfish,file='/Users/kirstenelstner/Desktop/finfish_model_lengths.csv')
```




Model Version 2 - Incorporating Bias (data z-scored)

```{r}
# make data list with z-scored data
datZ <- list(
  B = zScore(finfish$Angler_Hand_Measurement_TL_mm),
  O = zScore(finfish$Algorithm_Estimate_mm), 
  A = zScore(finfish$Adjusted_Estimate_mm),
  FishID = finfish$FishID.b,  
  AnglerID = finfish$AnglerNumber
)

dat <- list(
  B = finfish$Angler_Hand_Measurement_TL_mm,
  O = finfish$Algorithm_Estimate_mm, 
  A = finfish$Adjusted_Estimate_mm,
  FishID = finfish$FishID.b,  
  AnglerID = finfish$AnglerNumber
)


# fit model
M2 <- ulam(
  alist(
    TrueSize[FishID] ~ dnorm(0, 400),
    B ~ dnorm(TrueSize[FishID], Bsig), # board measurements are distributed with some bias around the true measurement of a fish 
    O ~ dnorm(TrueSize[FishID] + O_Bias, Osig), # original Poseidon estimates are distributed with some bias around the true measurement of a fish
    A ~ dnorm(TrueSize[FishID] + A_Bias, Asig), # adjusted Poseidon estimates are distributed with some bias around the true measurement of a fish
    
    # define bias parameters
    O_Bias ~ dnorm(0, 200), 
    A_Bias ~ dnorm(0, 200), 
    
    # variance priors
    Bsig ~ dexp(1),
    Osig ~ dexp(1),
    Asig ~ dexp(1)
  ), data = dat, chains = 4 
)


# sample from the posterior
PostM2 <- extract.samples(M2, n = 10000)


# extract bias terms across all Anglers and convert from z-score to mm
M2.bias <- data.frame(
  PostM2$O_Bias,
  PostM2$A_Bias
)

# calculate MAP and 95% HPDI for bias terms 

# MAP
round(chainmode(PostM2$O_Bias), digits = 5)*sd(finfish$Angler_Hand_Measurement_TL_mm) # O_bias = -0.108 mm
round(chainmode(PostM2$A_Bias), digits = 5)*sd(finfish$Angler_Hand_Measurement_TL_mm) # A_bias = 0.139 mm


# 95% HPDI
round(HPDI(PostM2$O_Bias, prob = 0.95), digits = 5)*sd(finfish$Angler_Hand_Measurement_TL_mm) # O_Bias = [-4.941  5.270]
round(HPDI(PostM2$A_Bias, prob = 0.95), digits = 5)*sd(finfish$Angler_Hand_Measurement_TL_mm) # BiasHat_A = [-2.153  2.317]


# Bias plot
Names <- c("Original Estimate", "Adjusted Estimate")
names(M2.bias) <- Names
M2.bias <- melt(M2.bias)
names2 <- c("Measurement", "Bias")
names(M2.bias) <- names2
# make plot that overlays posterior densities of bias estimates
cols <- c("gray", "red")
BiasedPlot = ggplot(data = M2.bias, aes(x=Bias, fill=Measurement)) +
  geom_density(alpha=.25) +
  labs(y= "Density", x = "Posterior Estimate of Measurement Bias (mm)") +
  scale_fill_manual(values = cols) +
  theme_bw()
  
BiasedPlot + theme(plot.title = element_text(hjust = 0.5))


# Variance Plot
M2.var <- data.frame(
  PostM2$Osig,
  PostM2$Asig,
  PostM2$Bsig
)

Names <- c("Original Estimate", "Adjusted Estimate", "Hand Measurement")
names(M2.var) <- Names
M2.var <- melt(M2.var)
names2 <- c("Measurement", "Variance")
names(M2.var) <- names2

cols <- c("gray", "red", "lightblue")

# create density plots of variance estimates across fisheries
M2.var_plot = ggplot(data = M2.var, aes(x=Variance, fill=Measurement)) +
  geom_density(alpha=0.5) +
  labs(y= "Density", x = "Posterior Estimates of Measurement Variance") +
  scale_fill_manual(values = cols) + 
  theme_bw()

var.plot <- M2.var_plot + theme(plot.title = element_text(hjust = 0.5))
var.plot
VarBias_Plot <- ggarrange(M2.var_plot, BiasedPlot, nrow = 2, ncol = 1)

VarBias_Plot
```


Posterior Predictive Check with FishID.b # 16 (TagID 20792)
```{r}
# sample from M2 posterior
PostM2 <- extract.samples(M2, n = 10000)

str(PostM2)

# draw samples from true size
TS.sim <- PostM2$TrueSize[, 16]
# draw samples from O_Bias Estimate
OBias.sim <- PostM2$O_Bias
# draw samples from Osig
Osig.sim <- PostM2$Osig
drawsFunc <- function(x) {
  result <- rnorm(1, 0, x)
  return(result)
}
ObiasRandomDraws <- apply(Osig.sim, 1, drawsFunc)
simDat.Original <- TS.sim + OBias.sim + ObiasRandomDraws
Fish16 <- subset(finfish, FishID.b==16)

plot(hist(zScore(Fish16$Algorithm_Estimate_mm), breaks=3), xlim = c(-3, 3), ylim = c(0, 5))
lines(density(simDat.Original), col="red", xlim = c(-3, 3), ylim = c(0, 5))
```



test M2 with simulated data
```{r}
# simulate original algorithm, adjusted, and board measurements
Original <- rnorm(299, 300, 80) + rnorm(299, 50, 10)
Adjusted <- rnorm(299, 300, 70) + rnorm(299, 30, 10)
Board <- rnorm(299, 300, 50)

OBias <- rnorm(299, 50, 10)
ABias <- rnorm(299, 30, 10)
BBias <- rnorm(299, 0, 5)

plot(density(OBias), col = "black", xlim = c(-50, 100), ylim=c(0,0.2))
lines(density(ABias), col = "red", xlim = c(-50, 100), ylim=c(0,0.2))
lines(density(BBias), col = "blue", xlim = c(-50, 100), ylim=c(0,0.2))
legend("topright", c("Original Bias", "Adjusted Bias", "Board Bias"),
       col =c("black","red", "blue"), lty=1)


# first z-score these measurements then put them into their own list
simDat <- list(
  simB = zScore(Board),
  simO = zScore(Original),
  simA = zScore(Adjusted),
  FishID = finfish$FishID.b,
  AnglerID = finfish$AnglerID
)

# fit model using simulated data
M2S <- ulam(
  alist(
    TrueSize[FishID] ~ dnorm(0, 2),
    simB ~ dnorm(TrueSize[FishID], Bsig), 
    simO ~ dnorm(TrueSize[FishID] + O_Bias, Osig), 
    simA ~ dnorm(TrueSize[FishID] + A_Bias, Asig), 
    O_Bias ~ dnorm(0, 2), 
    A_Bias ~ dnorm(0, 2), 
    Bsig ~ dexp(1),
    Osig ~ dexp(1),
    Asig ~ dexp(1)
  ), data = simDat, chains = 4 
)

# sample from posterior
simPost <- extract.samples(M2S, n = 10000)

OBias.post <- simPost$O_Bias
ABias.post <- simPost$A_Bias

OBias.prior <-rnorm(1000, 0, 2)
  
plot(density(zScore(OBias)), col = "black", ylim = c(0,5), main = "Original Bias Estimates")
lines(density(OBias.prior), col = "blue", ylim = c(0,5))
lines(density(OBias.post), col = "red", ylim = c(0,5))
legend("topright", c("Raw Data", "Prior", "Posterior"),
       col =c("black", "blue", "red"), lty=1)


round(chainmode(simPost$O_Bias), digits = 3)*sd(Board) # O_Bias = -0.018
round(chainmode(simPost$A_Bias), digits = 3)*sd(Board) # A_Bias = -0.004




```




All Fisheries Models: 

This first model characterizes Total Error across all fisheries 

```{r}
# import data for all fisheries and only include yes data
abalone <- read.csv("~/Desktop/SIO/Research/Poseidon Project/Final Analysis/Final Data/Fishery-Specific Data/csv/AbaloneCleanedJanuary2023.csv")
abalone <- subset(abalone, Include == "yes")

# this finfish data has been updated to incorporate modeled lengths 
finfish <- read.csv("~/Desktop/SIO/Research/Poseidon Project/Final Analysis/Final Data/finfish_model_lengths.csv")

lobster <- read.csv("~/Desktop/SIO/Research/Poseidon Project/Final Analysis/Final Data/Fishery-Specific Data/csv/CaliforniaSpinyLobster_Cleaned_Jan2023.csv")
lobster <- subset(lobster, Include == "yes")

penshell <- read.csv("~/Desktop/SIO/Research/Poseidon Project/Final Analysis/Final Data/Fishery-Specific Data/csv/Penshell_Cleaned_January2023.csv")
penshell <- subset(penshell, Included == "yes")
penshell <- subset(penshell, penshell$Algorithm_Estimate_mm < 500)

# make list
all_fisheries_list <- list(
  AbaloneOD = abalone$Algorithm_Total_Error_mm,
  AbaloneAD = abalone$Adjusted_Total_Error_mm,
  FinfishOD = finfish$New_Algorithm_Total_Error_mm,
  FinfishAD = finfish$New_Adjusted_Total_Error_mm,
  LobsterOD = lobster$Algorithm_Total_Error_mm,
  LobsterAD = lobster$Adjusted_Total_Error_mm,
  PenshellOD = penshell$Algorithm_Total_Error_mm,
  PenshellAD = penshell$Adjusted_Total_Error_mm)


model.all <- ulam(
  alist(
    # abalone
    AbaloneOD ~ dnorm(mu_AOD, sig_AOD),
    AbaloneAD ~ dnorm(mu_AAD, sig_AAD),
    mu_AOD ~ dnorm(0, 2),
    mu_AAD ~ dnorm(0, 2),
    sig_AOD ~ dexp(1),
    sig_AAD ~ dexp(1),
    
    # finfish
    FinfishOD ~ dnorm(mu_FOD, sig_FOD),
    FinfishAD ~ dnorm(mu_FAD, sig_FAD),
    mu_FOD ~ dnorm(0, 2),
    mu_FAD ~ dnorm(0, 2),
    sig_FOD ~ dexp(1),
    sig_FAD ~ dexp(1),
    
    # lobster
    LobsterOD ~ dnorm(mu_LOD, sig_LOD),
    LobsterAD ~ dnorm(mu_LAD, sig_LAD),
    mu_LOD ~ dnorm(0, 2),
    mu_LAD ~ dnorm(0, 2),
    sig_LOD ~ dexp(1),
    sig_LAD ~ dexp(1),
    
    # penshell
    PenshellOD ~ dnorm(mu_POD, sig_POD),
    PenshellAD ~ dnorm(mu_PAD, sig_PAD),
    mu_POD ~ dnorm(0, 2),
    mu_PAD ~ dnorm(0, 2),
    sig_POD ~ dexp(1),
    sig_PAD ~ dexp(1)
    
  ), data = all_fisheries_list, chains = 4)


# extract samples from posterior
Post.all <- extract.samples(model.all, n=10000)



# calculate MAP and 95% HPDI for each error parameter estimate

# abalone MAP
round(chainmode(Post.all$mu_AOD), digits = 2) # mu_AOD = 13.92
round(chainmode(Post.all$mu_AAD), digits = 2) # mu_AAD = -1.48
# abalone 95% HPDI
round(HPDI(Post.all$mu_AOD, prob = 0.95), digits = 2) # mu_AOD = [11.54, 16.52] 
round(HPDI(Post.all$mu_AAD, prob = 0.95), digits = 2) # mu_AOD = [-2.58, -0.51]


# finfish MAP
round(chainmode(Post.all$mu_FOD), digits = 2) # mu_FOD = 8.09
round(chainmode(Post.all$mu_FAD), digits = 2) # mu_FAD = 24.94
# finfish 95% HPDI
round(HPDI(Post.all$mu_FOD, prob = 0.95), digits = 2) # mu_FOD = [4.82, 11.15 ] 
round(HPDI(Post.all$mu_FAD, prob = 0.95), digits = 2) # mu_FOD = [22.08, 27.19]


# lobster MAP
round(chainmode(Post.all$mu_LOD), digits = 2) # mu_LOD = 14.23
round(chainmode(Post.all$mu_LAD), digits = 2) # mu_LAD = 1.63
# lobster 95% HPDI
round(HPDI(Post.all$mu_LOD, prob = 0.95), digits = 2) # mu_LOD = [12.21, 15.74] 
round(HPDI(Post.all$mu_LAD, prob = 0.95), digits = 2) # mu_LOD = [1.05, 2.29]


# penshell MAP
round(chainmode(Post.all$mu_POD), digits = 2) # mu_POD = 3.42
round(chainmode(Post.all$mu_PAD), digits = 2) # mu_PAD = 1.48
# lobster 95% HPDI
round(HPDI(Post.all$mu_POD, prob = 0.95), digits = 2) # mu_POD = [2.95, 4.10] 
round(HPDI(Post.all$mu_PAD, prob = 0.95), digits = 2) # mu_POD = [1.02, 1.93]
```



this second model is the same as the first model but examines percent error


```{r}
# import data for all fisheries and only include yes data
abalone <- read.csv("~/Desktop/SIO/Research/Poseidon Project/Final Analysis/Final Data/Fishery-Specific Data/csv/AbaloneCleanedJanuary2023.csv")
abalone <- subset(abalone, Include == "yes")

# this finfish data has been updated to incorporate modeled lengths 
finfish <- read.csv("~/Desktop/SIO/Research/Poseidon Project/Final Analysis/Final Data/finfish_model_lengths.csv")

lobster <- read.csv("~/Desktop/SIO/Research/Poseidon Project/Final Analysis/Final Data/Fishery-Specific Data/csv/CaliforniaSpinyLobster_Cleaned_Jan2023.csv")
lobster <- subset(lobster, Include == "yes")

penshell <- read.csv("~/Desktop/SIO/Research/Poseidon Project/Final Analysis/Final Data/Fishery-Specific Data/csv/Penshell_Cleaned_January2023.csv")
penshell <- subset(penshell, Included == "yes")
penshell <- subset(penshell, penshell$Algorithm_Estimate_mm < 500)

# make list
all_fisheries_list.percent <- list(
  AbaloneOD = abalone$Algorithm_Percent_Error,
  AbaloneAD = abalone$Adjusted_Percent_Error,
  FinfishOD = finfish$New_Algorithm_Percent_Error,
  FinfishAD = finfish$New_Adjusted_Percent_Error,
  LobsterOD = lobster$Algorithm_Percent_Error,
  LobsterAD = lobster$Adjusted_Percent_Error,
  PenshellOD = penshell$Algorithm_Percent_Error,
  PenshellAD = penshell$Adjusted_Percent_Error)


model.all <- ulam(
  alist(
    # abalone
    AbaloneOD ~ dnorm(mu_AOD, sig_AOD),
    AbaloneAD ~ dnorm(mu_AAD, sig_AAD),
    mu_AOD ~ dnorm(0, 2),
    mu_AAD ~ dnorm(0, 2),
    sig_AOD ~ dexp(1),
    sig_AAD ~ dexp(1),
    
    # finfish
    FinfishOD ~ dnorm(mu_FOD, sig_FOD),
    FinfishAD ~ dnorm(mu_FAD, sig_FAD),
    mu_FOD ~ dnorm(0, 2),
    mu_FAD ~ dnorm(0, 2),
    sig_FOD ~ dexp(1),
    sig_FAD ~ dexp(1),
    
    # lobster
    LobsterOD ~ dnorm(mu_LOD, sig_LOD),
    LobsterAD ~ dnorm(mu_LAD, sig_LAD),
    mu_LOD ~ dnorm(0, 2),
    mu_LAD ~ dnorm(0, 2),
    sig_LOD ~ dexp(1),
    sig_LAD ~ dexp(1),
    
    # penshell
    PenshellOD ~ dnorm(mu_POD, sig_POD),
    PenshellAD ~ dnorm(mu_PAD, sig_PAD),
    mu_POD ~ dnorm(0, 2),
    mu_PAD ~ dnorm(0, 2),
    sig_POD ~ dexp(1),
    sig_PAD ~ dexp(1)
    
  ), data = all_fisheries_list.percent, chains = 4)


# extract samples from posterior
Post.all <- extract.samples(model.all, n=10000)



# calculate MAP and 95% HPDI for each error parameter estimate

# abalone MAP
round(chainmode(Post.all$mu_AOD), digits = 2) # mu_AOD = 9.28%
round(chainmode(Post.all$mu_AAD), digits = 2) # mu_AAD = -0.60%



# finfish MAP
round(chainmode(Post.all$mu_FOD), digits = 2) # mu_FOD = 7.1%
round(chainmode(Post.all$mu_FAD), digits = 2) # mu_FAD = 11.58%



# lobster MAP
round(chainmode(Post.all$mu_LOD), digits = 2) # mu_LOD = 14.52%
round(chainmode(Post.all$mu_LAD), digits = 2) # mu_LAD = 1.58%



# penshell MAP
round(chainmode(Post.all$mu_POD), digits = 2) # mu_POD = 2.89%
round(chainmode(Post.all$mu_PAD), digits = 2) # mu_PAD = 1.3%

```

Create violin plot that displays length estimate error across fisheries for both total error and percent error

```{r}
library("vioplot")

penshell.edited <- subset(penshell, penshell$Algorithm_Estimate_mm < 500)

colors <- c("gray", "red", "gray", "red", "gray", "red", "gray", "red")



# total error plot
TE.violin.plot <- vioplot(
  abalone$Algorithm_Total_Error_mm,
  abalone$Adjusted_Total_Error_mm,
  finfish$New_Algorithm_Total_Error_mm, 
  finfish$New_Adjusted_Total_Error_mm, 
  lobster$Algorithm_Total_Error_mm,
  lobster$Adjusted_Total_Error_mm,
  penshell.edited$Algorithm_Total_Error_mm,
  penshell.edited$Adjusted_Total_Error_mm,
  col = colors,
  xlab = "Fishery",
  ylab = "Total Error (mm)",
  ylim = c(-300, 200),
  xaxt = "n"
  )

mtext("Abalone", side = 1, at = 1.5, line = 1)
mtext("Finfish", side = 1, at = 3.5, line = 1)
mtext("Lobster", side = 1, at = 5.5, line = 1)
mtext("Penshell", side = 1, at = 7.5, line = 1)
mtext("A", side = 3, at = 1, line = 1, cex =2)

legend("bottomright", legend = c("Original Software Estimate", "User-Adjusted Estimate"), fill = c("gray", "red"), cex = 1)

abline(a=0, b=0, lty=2)

```

```{r}
# percent error plot
PE.violin.plot <- vioplot(
  abalone$Algorithm_Percent_Error,
  abalone$Adjusted_Percent_Error,
  finfish$New_Algorithm_Percent_Error, 
  finfish$New_Adjusted_Percent_Error, 
  lobster$Algorithm_Percent_Error,
  lobster$Adjusted_Percent_Error,
  penshell.edited$Algorithm_Percent_Error,
  penshell.edited$Adjusted_Percent_Error,
  col = colors,
  xlab = "Fishery",
  ylab = "Percent Error (%)",
  ylim = c(-150, 160),
  xaxt = "n"
  )

mtext("Abalone", side = 1, at = 1.5, line = 1)
mtext("Finfish", side = 1, at = 3.5, line = 1)
mtext("Lobster", side = 1, at = 5.5, line = 1)
mtext("Penshell", side = 1, at = 7.5, line = 1)
mtext("B", side = 3, at = 1, line = 1, cex =2)

legend("bottomright", legend = c("Original Software Estimate", "User-Adjusted Estimate"), fill = c("gray", "red"), cex = 1)

abline(a=0, b=0, lty=2)
```


Create violin plots that display total error across users for abalone, finfish and penshell

abalone
```{r}
library(vioplot)
library(ggplot2)

ao <- ggplot(abalone, aes(x=User_ID, y=Algorithm_Total_Error_mm, group = User_ID)) + 
  geom_violin(col="gray", fill="gray") +
  coord_cartesian(ylim=c(-250, 175)) +
  xlab("Software User ID") +
  ylab("Total Error (mm)")+
  ggtitle("Abalone (n = 13)") +
  geom_hline(yintercept=0,linetype=2) +
  theme_bw()

aop <- ao + geom_boxplot(width=0.2, outlier.shape = NA) + theme(axis.text.x = element_blank(),
           axis.ticks.x = element_blank())

aop
```

```{r}
aa <- ggplot(abalone, aes(x=User_ID, y=Adjusted_Total_Error_mm, group = User_ID)) + 
  geom_violin(col="red", fill="red") +
  coord_cartesian(ylim=c(-250, 175)) +
  xlab("Software User ID") +
  ylab("Total Error (mm)")+
  #ggtitle("Abalone") +
  geom_hline(yintercept=0,linetype=2) +
  theme_bw()

aap <- aa + geom_boxplot(width=0.2, outlier.shape = NA) + theme(axis.text.x = element_blank(),
           axis.ticks.x = element_blank())

aap
```

finfish
```{r}

fo <- ggplot(finfish, aes(x=AnglerNumber, y=New_Algorithm_Total_Error_mm, group = AnglerNumber)) + 
  geom_violin(col="gray", fill="gray") +
  coord_cartesian(ylim=c(-250, 175)) +
  xlab("Software User ID") +
  ylab("Total Error (mm)")+
  ggtitle("Finfish (n = 9)") +
  geom_hline(yintercept=0,linetype=2) +
  theme_bw()

fop <- fo + geom_boxplot(width=0.2, outlier.shape = NA) + theme(axis.text.x = element_blank(),
           axis.ticks.x = element_blank())

fop
```


```{r}
fa <- ggplot(finfish, aes(x=AnglerNumber, y=New_Adjusted_Total_Error_mm, group = AnglerNumber)) + 
  geom_violin(col="red", fill="red") +
  coord_cartesian(ylim=c(-250, 175)) +
  xlab("Software User ID") +
  ylab("Total Error (mm)")+
  #ggtitle("Finfish") +
  geom_hline(yintercept=0,linetype=2) +
  theme_bw()

fap <- fa + geom_boxplot(width=0.2, outlier.shape = NA) + theme(axis.text.x = element_blank(),
           axis.ticks.x = element_blank())

fap

```


penshell

```{r}
penshell.edited <- subset(penshell, penshell$Algorithm_Estimate_mm < 500)

po <- ggplot(penshell.edited, aes(x=User_ID, y=Algorithm_Total_Error_mm, group = User_ID)) + 
  geom_violin(col="gray", fill="gray") +
  coord_cartesian(ylim=c(-250, 175)) +
  xlab("Software User ID") +
  ylab("Total Error (mm)")+
  ggtitle("Pen Shell (n = 25)") +
  geom_hline(yintercept=0,linetype=2) +
  theme_bw()

pop <- po + geom_boxplot(width=0.2, outlier.shape = NA) + theme(axis.text.x = element_blank(),
           axis.ticks.x = element_blank())

pop

```


```{r}
penshell.edited <- subset(penshell, penshell$Algorithm_Estimate_mm < 500)

pa <- ggplot(penshell.edited, aes(x=User_ID, y=Adjusted_Total_Error_mm, group = User_ID)) + 
  geom_violin(col="red", fill="red") +
  coord_cartesian(ylim=c(-250, 175)) +
  xlab("Software User ID") +
  ylab("Total Error (mm)")+
  #ggtitle("Penshell") +
  geom_hline(yintercept=0,linetype=2) +
  theme_bw()

pap <- pa + geom_boxplot(width=0.2, outlier.shape = NA) + theme(axis.text.x = element_blank(),
           axis.ticks.x = element_blank())

pap 
```


```{r}
library(ggpubr)

user_figure <- ggarrange(aop, fop, pop,
                         aap, fap, pap,
          nrow=2,
          ncol=3)

user_figure

annotate_figure(user_figure, top = text_grob("Size Estimate Variability Across Users", 
               face = "bold", size = 14))
```




For finfish, create violin plot that shows how original/adjusted Poseidon Total Error compares to Angler Total Error
```{r}
library(vioplot)
library(reshape)
library(reshape2)
library(ggplot2)

finfish$Angler_Total_Error_mm <- finfish$Angler_Hand_Measurement_TL_mm - finfish$Scientist_Hand_Measured_TL_mm

vioplot(
  finfish$Algorithm_Total_Error_mm,
  finfish$Adjusted_Total_Error_mm,
  finfish$Angler_Total_Error_mm
)


TE.data <- data.frame(
  finfish$Algorithm_Total_Error_mm,
  finfish$Adjusted_Total_Error_mm,
  finfish$Angler_Total_Error_mm)
  
  
  # get data into required format for plotting
Names <- c("Original Estimate", "Adjusted Estimate", "Hand Measurement")
names(TE.data) <- Names
TE.data <- melt(TE.data)
names2 <- c("MeasurementType", "TotalError")
names(TE.data) <- names2


# make violoin plot that compares original, adjusted and hand measurements
cols <- c("gray", "red", "lightblue")




ComparisonPlot <- ggplot(TE.data, aes(x=MeasurementType, y=TotalError, fill = MeasurementType)) + 
  geom_violin(alpha=0.5) +
  xlab("Length Measurement Type") +
  ylab("Total Error (mm)") +
  geom_hline(yintercept=0,linetype=2) +
  theme_bw()

  
plot <- ComparisonPlot + 
  scale_fill_manual(values=c("darkgray", "red", "lightblue")) +
  theme(legend.title=element_blank())
plot

library(ggpubr)
ggarrange(plot, var.plot, nrow = 2, ncol = 1)

figure
```
